### 2020-05-08 inst/doc/ms-perks-features.Rmd
version 1.1.0

---
title: "modelStudio - perks and features"
author: "Hubert Baniecki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modelStudio - perks and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

`modelStudio::modelStudio` computes various (instance and dataset level) model explanations and produces
an interactive, customisable dashboard. It consists of multiple panels for plots with their short descriptions.
Easily save and share the HTML dashboard with others. Tools for model exploration unite with tools for
EDA (Exploratory Data Analysis) to give a broad overview of the model behavior.

Let's use `HR` dataset to explore `modelStudio` parameters:

```{r results="hide"}
train <- DALEX::HR
train$fired <- as.factor(ifelse(train$status == "fired", 1, 0))
train$status <- NULL

head(train)
```

```{r echo = FALSE, fig.align='center'}
knitr::kable(head(train), digits = 2, caption = "DALEX::HR dataset")
```

Prepare `HR_test` data and a `ranger` model for the explainer:

```{r results="hide", eval = FALSE}
# fit a ranger model
library("ranger")
model <- ranger(fired ~., data = train, probability = TRUE)

# prepare validation dataset
test <- DALEX::HR_test[1:1000,]
test$fired <- ifelse(test$status == "fired", 1, 0)
test$status <- NULL

# create an explainer for the model
explainer <- DALEX::explain(model,
                            data = test,
                            y = test$fired)

# start modelStudio
library("modelStudio")
```

-------------------------------------------------------------------

## modelStudio parameters

### instance explanations

Pass data points to the `new_observation` parameter for instance explanations
such as [Break Down](https://pbiecek.github.io/ema/breakDown.html),
[Shapley Values](https://pbiecek.github.io/ema/shapley.html) and
[Ceteris Paribus](https://pbiecek.github.io/ema/ceterisParibus.html) Profiles.
Use `new_observation_y` to show their true labels.

```{r eval = FALSE}
new_observation <- test[1:3,]
rownames(new_observation) <- c("John Snow", "Arya Stark", "Samwell Tarly")
true_labels <- test[1:3,]$fired

modelStudio(explainer,
            new_observation = new_observation,
            new_observation_y  = true_labels)
```

### grid size

Achieve bigger or smaller `modelStudio` grid with `facet_dim` parameter.

```{r eval = FALSE}
# small dashboard with 2 panels
modelStudio(explainer,
            facet_dim = c(1,2))

# large dashboard with 9 panels
modelStudio(explainer,
            facet_dim = c(3,3))
```

### animations

Manipulate `time` parameter to set animation length. Value 0 will make
them invisible.

```{r eval = FALSE}
# slow down animations
modelStudio(explainer,
            time = 1000)

# turn off animations
modelStudio(explainer,
            time = 0)
```

### more calculations means more time

`N` is a number of observations used for calculation of
[Partial Dependence](https://pbiecek.github.io/ema/partialDependenceProfiles.html)
and [Accumulated Dependence](https://pbiecek.github.io/ema/accumulatedLocalProfiles.html) Profiles. 
`10*N` is a number of observations used for calculation of
[Feature Importance](https://pbiecek.github.io/ema/featureImportance.html).
`B` is a number of permutation rounds used for calculation of
[Shapley Values](https://pbiecek.github.io/ema/shapley.html) and
[Feature Importance](https://pbiecek.github.io/ema/featureImportance.html).
Decrease `N` and `B` parameters to lower the computation time or increase
them to get more accurate empirical results.

```{r eval = FALSE}
# faster, less precise
modelStudio(explainer,
            N = 200, B = 5)

# slower, more precise
modelStudio(explainer,
            N = 500, B = 15)
```

### no EDA mode

Don't compute the EDA plots if they are not needed. Set the `eda` parameter to `FALSE`.

```{r eval = FALSE}
modelStudio(explainer,
            eda = FALSE)
```

### progress bar

Hide computation progress bar messages with `show_info` parameter.

```{r eval = FALSE}
modelStudio(explainer,
            show_info = FALSE)
```

### viewer or browser?

Change `viewer` parameter to set where to display `modelStudio`.
Best described in `r2d3` documentation:
[`r2d3 viewer argument`](https://rstudio.github.io/r2d3/articles/visualization_options.html#viewer).

```{r eval = FALSE}
modelStudio(explainer,
            viewer = "browser")
```

-------------------------------------------------------------------

## parallel computation

Speed up `modelStudio` computation by setting `parallel` parameter to `TRUE`.
It uses [`parallelMap`](https://www.rdocumentation.org/packages/parallelMap) package
to calculate local explainers faster. It is really useful when using `modelStudio` with
complicated models, vast datasets or **many observations are being processed**.

All options can be set outside of function call.
More on that [here](https://github.com/berndbischl/parallelMap#being-lazy-configuration).

```{r eval = FALSE}
# set up the cluster
options(
  parallelMap.default.mode        = "socket",
  parallelMap.default.cpus        = 4,
  parallelMap.default.show.info   = FALSE
)

# calculations of local explanations will be distributed into 4 cores
modelStudio(explainer,
            new_observation = test[1:16,],
            parallel = TRUE)
```

--------------------------------------------------------------------

## additional options

Customize some of `modelStudio` looks by overwriting default options returned
by the `ms_options()` function. Full list of options:
[documentation](http://modelstudio.drwhy.ai/reference/ms_options.html).

```{r eval = FALSE}
# set additional graphical parameters
new_options <- ms_options(
  show_subtitle = TRUE,
  bd_subtitle = "Hello World",
  line_size = 5,
  point_size = 9,
  line_color = "pink",
  point_color = "purple",
  bd_positive_color = "yellow",
  bd_negative_color = "orange"
)

modelStudio(explainer,
            options = new_options)
```

All visual options can be changed after the calculations using `ms_update_options()`.

```{r eval = FALSE}
old_ms <- modelStudio(explainer)
old_ms

# update the options
new_ms <- ms_update_options(old_ms,
                            time = 0,
                            facet_dim = c(1,2),
                            margin_left = 150)
new_ms
```

-------------------------------------------------------------------

## update observations

Use `ms_update_observations()` to add more observations with their local explanations to the `modelStudio`.

```{r eval = FALSE}
old_ms <- modelStudio(explainer)
old_ms

# add new observations
plus_ms <- ms_update_observations(old_ms,
                                  explainer,
                                  new_observation = test[101:102,])
plus_ms

# overwrite old observations
new_ms <- ms_update_observations(old_ms,
                                 explainer,
                                 new_observation = test[103:104,],
                                 overwrite = TRUE)
new_ms
```

## DALEXtra

Use `explain_*()` functions from the [DALEXtra](https://github.com/ModelOriented/DALEXtra)
package to explain various models. Bellow basic example of making `modelStudio` for
a `mlr` model using `explain_mlr()`.

```{r eval = FALSE}
library(DALEXtra)
library(mlr)

# fit a model
task <- makeClassifTask(id = "task", data = train, target = "fired")
learner <- makeLearner("classif.ranger", predict.type = "prob")
model <- train(learner, task)

# create an explainer for the model
explainer_mlr <- explain_mlr(model,
                             data = test,
                             y = test$fired,
                             label = "mlr")

# make a studio for the model
modelStudio(explainer_mlr)
```

## References

* Theoretical introduction to the plots: [Explanatory Model Analysis. Explore, Explain and Examine Predictive Models.](https://pbiecek.github.io/ema)
* The input object is implemented in [DALEX](https://modeloriented.github.io/DALEX/)
* Feature Importance, Ceteris Paribus, Partial Dependence and Accumulated Dependence plots
are implemented in [ingredients](https://modeloriented.github.io/ingredients/)
* Break Down and Shapley Values plots are implemented in [iBreakDown](https://modeloriented.github.io/iBreakDown/)