### 2021-01-07 inst/doc/ms-perks-features.Rmd
version 2.1.1

---
title: "modelStudio - perks and features"
author: "Hubert Baniecki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modelStudio - perks and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

The `modelStudio()` function computes various (instance and dataset level) model explanations 
and produces a customisable dashboard, which consists of multiple panels for plots with their
short descriptions. Easily save the dashboard and share it with others. Tools for
[Explanatory Model Analysis](https://ema.drwhy.ai/) unite with tools for Exploratory Data Analysis 
to give a broad overview of the model behavior.

Let's use `HR` dataset to explore `modelStudio` parameters:

```{r results="hide"}
train <- DALEX::HR
train$fired <- as.factor(ifelse(train$status == "fired", 1, 0))
train$status <- NULL

head(train)
```

```{r echo = FALSE, fig.align='center'}
knitr::kable(head(train), digits = 2, caption = "DALEX::HR dataset")
```

Prepare `HR_test` data and a `ranger` model for the explainer:

```{r results="hide", eval = FALSE}
# fit a ranger model
library("ranger")
model <- ranger(fired ~., data = train, probability = TRUE)

# prepare validation dataset
test <- DALEX::HR_test[1:1000,]
test$fired <- ifelse(test$status == "fired", 1, 0)
test$status <- NULL

# create an explainer for the model
explainer <- DALEX::explain(model,
                            data = test,
                            y = test$fired)

# start modelStudio
library("modelStudio")
```

-------------------------------------------------------------------

## modelStudio parameters

### instance explanations

Pass data points to the `new_observation` parameter for instance explanations
such as [Break Down](https://ema.drwhy.ai/breakDown.html),
[Shapley Values](https://ema.drwhy.ai/shapley.html) and
[Ceteris Paribus](https://ema.drwhy.ai/ceterisParibus.html) Profiles.
Use `new_observation_y` to show their true labels.

```{r eval = FALSE}
new_observation <- test[1:3,]
rownames(new_observation) <- c("John Snow", "Arya Stark", "Samwell Tarly")
true_labels <- test[1:3,]$fired

modelStudio(explainer,
            new_observation = new_observation,
            new_observation_y  = true_labels)
```

If `new_observation = NULL`, then choose `new_observation_n` observations, evenly spread by the order of `y_hat`. This shall always include the observations, which ids are `which.min(y_hat)` and `which.max(y_hat)`.

```{r eval = FALSE}
modelStudio(explainer, new_observation_n = 5) # default is 3
```

### grid size

Achieve bigger or smaller `modelStudio` grid with `facet_dim` parameter.

```{r eval = FALSE}
# small dashboard with 2 panels
modelStudio(explainer,
            facet_dim = c(1,2))

# large dashboard with 9 panels
modelStudio(explainer,
            facet_dim = c(3,3))
```

### animations

Manipulate `time` parameter to set animation length. Value 0 will make
them invisible.

```{r eval = FALSE}
# slow down animations
modelStudio(explainer,
            time = 1000)

# turn off animations
modelStudio(explainer,
            time = 0)
```

### more calculations means more time

- `N` is a number of observations used for calculation of
[Partial Dependence](https://ema.drwhy.ai/partialDependenceProfiles.html)
and [Accumulated Dependence](https://ema.drwhy.ai/accumulatedLocalProfiles.html) Profiles (default is `300`). 
- `N_fi` is a number of observations used for calculation of
[Feature Importance](https://ema.drwhy.ai/featureImportance.html) (default is `10*N`).
- `B` is a number of permutation rounds used for calculation of
[Shapley Values](https://ema.drwhy.ai/shapley.html) (default is `10`).
- `B_fi` is a number of permutation rounds used for calculation of
[Feature Importance](https://ema.drwhy.ai/featureImportance.html) (default is `B`).

Decrease `N` and `B` parameters to lower the computation time or increase
them to get more accurate empirical results.

```{r eval = FALSE}
# faster, less precise
modelStudio(explainer,
            N = 200, B = 5)

# slower, more precise
modelStudio(explainer,
            N = 500, B = 15)
```

### no EDA mode

Don't compute the EDA plots if they are not needed. Set the `eda` parameter to `FALSE`.

```{r eval = FALSE}
modelStudio(explainer,
            eda = FALSE)
```

### progress bar

Hide computation progress bar messages with `show_info` parameter.

```{r eval = FALSE}
modelStudio(explainer,
            show_info = FALSE)
```

### viewer or browser?

Change `viewer` parameter to set where to display `modelStudio`.
[Best described in `r2d3` documentation](https://rstudio.github.io/r2d3/articles/visualization_options.html#viewer).

```{r eval = FALSE}
modelStudio(explainer,
            viewer = "browser")
```

-------------------------------------------------------------------

## parallel computation

Speed up `modelStudio` computation by setting `parallel` parameter to `TRUE`.
It uses [`parallelMap`](https://www.rdocumentation.org/packages/parallelMap) package
to calculate local explainers faster. It is really useful when using `modelStudio` with
complicated models, vast datasets or **many observations are being processed**.

All options can be set outside of the function call.
[How to use parallelMap](https://github.com/mlr-org/parallelMap#being-lazy-configuration).

```{r eval = FALSE}
# set up the cluster
options(
  parallelMap.default.mode        = "socket",
  parallelMap.default.cpus        = 4,
  parallelMap.default.show.info   = FALSE
)

# calculations of local explanations will be distributed into 4 cores
modelStudio(explainer,
            new_observation = test[1:16,],
            parallel = TRUE)
```

--------------------------------------------------------------------

## additional options

Customize some of the `modelStudio` looks by overwriting default options returned
by the `ms_options()` function.
[Full list of options](https://modelstudio.drwhy.ai/reference/ms_options.html).

```{r eval = FALSE}
# set additional graphical parameters
new_options <- ms_options(
  show_subtitle = TRUE,
  bd_subtitle = "Hello World",
  line_size = 5,
  point_size = 9,
  line_color = "pink",
  point_color = "purple",
  bd_positive_color = "yellow",
  bd_negative_color = "orange"
)

modelStudio(explainer,
            options = new_options)
```

All visual options can be changed after the calculations using `ms_update_options()`.

```{r eval = FALSE}
old_ms <- modelStudio(explainer)
old_ms

# update the options
new_ms <- ms_update_options(old_ms,
                            time = 0,
                            facet_dim = c(1,2),
                            margin_left = 150)
new_ms
```

-------------------------------------------------------------------

## update observations

Use `ms_update_observations()` to add more observations with their local explanations to the `modelStudio`.

```{r eval = FALSE}
old_ms <- modelStudio(explainer)
old_ms

# add new observations
plus_ms <- ms_update_observations(old_ms,
                                  explainer,
                                  new_observation = test[101:102,])
plus_ms

# overwrite old observations
new_ms <- ms_update_observations(old_ms,
                                 explainer,
                                 new_observation = test[103:104,],
                                 overwrite = TRUE)
new_ms
```

-------------------------------------------------------------------

## Shiny

Use the `widget_id` argument and `r2d3` package to render the `modelStudio` output in Shiny.
See [Using r2d3 with Shiny](https://rstudio.github.io/r2d3/articles/shiny.html) and consider 
the following example:

```{r eval = FALSE}
library(shiny)
library(r2d3)

ui <- fluidPage(
  textInput("text", h3("Text input"), 
            value = "Enter text..."),
  uiOutput('dashboard')
)

server <- function(input, output) {
  #:# id of div where modelStudio will appear
  WIDGET_ID = 'MODELSTUDIO'

  #:# create modelStudio 
  library(modelStudio)
  library(DALEX)
  model <- glm(survived ~., data = titanic_imputed, family = "binomial")
  explainer <- DALEX::explain(model,
                              data = titanic_imputed,
                              y = titanic_imputed$survived,
                              label = "Titanic GLM",
                              verbose = FALSE)
  ms <- modelStudio(explainer,
                    widget_id = WIDGET_ID,  #:# use the widget_id 
                    show_info = FALSE)    
  ms$elementId <- NULL                      #:# remove elementId to stop the warning

  #:# basic render d3 output
  output[[WIDGET_ID]] <- renderD3({
    ms
  })

  #:# use render ui to set proper width and height
  output$dashboard <- renderUI({
    d3Output(WIDGET_ID, width=ms$width, height=ms$height)
  })
}

shinyApp(ui = ui, server = server)
```

-------------------------------------------------------------------

## DALEXtra

Use `explain_*()` functions from the [DALEXtra](https://github.com/ModelOriented/DALEXtra/)
package to explain various models.

Bellow basic example of making `modelStudio` for a `mlr` model using `explain_mlr()`.

```{r eval = FALSE}
library(DALEXtra)
library(mlr)

# fit a model
task <- makeClassifTask(id = "task", data = train, target = "fired")
learner <- makeLearner("classif.ranger", predict.type = "prob")
model <- train(learner, task)

# create an explainer for the model
explainer_mlr <- explain_mlr(model,
                             data = test,
                             y = test$fired,
                             label = "mlr")

# make a studio for the model
modelStudio(explainer_mlr)
```

-------------------------------------------------------------------

## References

* Theoretical introduction to the plots: [Explanatory Model Analysis. Explore, Explain, and Examine Predictive Models.](https://ema.drwhy.ai/)
* The input object is implemented in [DALEX](https://modeloriented.github.io/DALEX/)
* Feature Importance, Ceteris Paribus, Partial Dependence and Accumulated Dependence explanations
are implemented in [ingredients](https://modeloriented.github.io/ingredients/)
* Break Down and Shapley Values explanations are implemented in [iBreakDown](https://modeloriented.github.io/iBreakDown/)